---
title: "Community vs SGD Regression"
author: "Danielle Barnas"
date: "2022-10-14"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# Install necessary libraries
```{r, eval = F}
if("tidyverse" %in% rownames(installed.packages()) == FALSE){install.packages("tidyverse")}
if("here" %in% rownames(installed.packages()) == FALSE){install.packages("here")}
if("ggrepel" %in% rownames(installed.packages()) == FALSE){install.packages("ggrepel")}
if("PNWColors" %in% rownames(installed.packages()) == FALSE){install.packages("PNWColors")}
if("vegan" %in% rownames(installed.packages()) == FALSE){install.packages("vegan")}
if("pairwiseAdonis" %in% rownames(installed.packages()) == FALSE){ devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")}
if("patchwork" %in% rownames(installed.packages()) == FALSE){install.packages("patchwork")}
if("ggmap" %in% rownames(installed.packages()) == FALSE){install.packages("ggmap")}
```


# Load Libraries
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(PNWColors)
library(vegan)
library(pairwiseAdonis)
library(patchwork)
library(ggmap)
library(wakefield)
```


# Read in data
```{r, message = F}

meta <- read_csv(here("Data", "Full_Metadata.csv"))

#comp <- read_csv(here("Data","Surveys","Species_Composition_2022.csv"))
taxa <- read_csv(here("Data","Surveys","Distinct_Taxa.csv"))
chem <- read_csv(here("Data","Biogeochem", "Nutrients_Processed_All.csv"))
turb <- read_csv(here("Data","Biogeochem", "July2022", "Turb_NC.csv"))
# richness, % richness of community pool, and % volume of community pool
Fric <- read_csv(here("Data", "Sp_FE_Vol.csv"))


# create color palette for plotting
mypalette <- pnw_palette(name="Bay")
midpalette <- pnw_palette(name = "Bay", n = 30)
largepalette <- pnw_palette(name = "Bay", n = 100)
```




# Process and Join data
```{r}
# reduce metadata
meta <- meta %>% 
  drop_na(lat) %>%  # removes Maya's project sites (not being used for my study)
  left_join(turb)


# join dfs
Full_data <- Fric %>% 
  left_join(meta) %>% 
  mutate(meanRugosity = if_else(CowTagID == "VSEEP", 0.97, meanRugosity)) %>%   # i'm a ninny and still haven't added the seep rugosity yet *sigh*
  left_join(chem) %>% 
  select(-Location)

myturb <- turb %>% 
  pivot_longer(cols = del15N:N_percent, names_to = "Parameters", values_to = "Values") %>% 
  right_join(Fric) %>% 
  select(-Location) %>% 
  relocate(Parameters, .after = Vol8D) %>% 
  relocate(Values, .after = Parameters) %>% 
  mutate(Season = "June2022")
  

```

# Visualize linear and nonlinear regressions: Richness and Volume
```{r}
# Plotting function
# create ggplot function
plotfun <- function(mydata = Full_data, x, y, myfacet, myformula) {

  x<-enquo(x)
  y<-enquo(y)

  plot <- ggplot(data = mydata,
                 aes(y = !!y,
                     x = !!x,
                     color = Season)) +
  geom_point() +
  geom_smooth(method = "lm", formula = myformula) +
  facet_wrap(enquo(myfacet), scales = "free") +
  labs(y = paste(as_label(y)),
       x = paste(as_label(x))) +
  theme_bw()+
  theme(legend.direction = "horizontal",
        legend.position = "top")
  
  return(plot)
}

# pvalue Functions: LM and Polynomial
pvalLM <- function(mydata = Full_data, myparam, myseason){
  
  Param_data <- mydata %>%
    select(CowTagID, NbSp:Vol8D, Season, Parameters, {{myparam}}) %>% 
    pivot_wider(names_from = Parameters, values_from = {{myparam}})
  
  ParamType <- as.character(myparam)
  
  p_df <- tibble(Parameter = as.character(),
                 ParamType = as.character(),
                 Dependent = as.character(),
                 Season = as.character(),
                 pvalue = as.numeric(),
                 r_squared = as.numeric(),
                 adj_r_squared = as.numeric())
  
  for(i in 2:6){ # 2:6 are our dependent variables NbSp, NbFEs, Vol8D
  
    for(j in 8:ncol(Param_data)){
      
      Param_data <- Param_data %>% 
        filter(Season == myseason)
      Season <- myseason
      
      Parameter <- colnames(Param_data)[j]
      Dependent <- colnames(Param_data)[i] # select dependent parameter
      
      if(is.na(Param_data[1,j])){ # some variables have no data in the wet season, so skip these variables
        p_df <- p_df
      } else {
        
      mod <- lm(paste(Dependent, "~", Parameter), data = Param_data)
      pvalue <- summary(mod)[4]$coefficients[8]
      r_squared <- summary(mod)[8]$r.squared
      adj_r_squared <- summary(mod)[9]$adj.r.squared

      temp <- as_tibble(cbind(Parameter, ParamType, Dependent, Season, 
                          pvalue, r_squared, adj_r_squared)) %>%
              mutate(pvalue = as.numeric(pvalue),
                     r_squared = as.numeric(r_squared),
                     adj_r_squared = as.numeric(adj_r_squared))

      p_df <- p_df %>%
              rbind(temp)
      }
    }
  }
return(p_df)
}


pvalpoly <- function(mydata = Full_data, myparam, myseason){
  
  Param_data <- mydata %>%
    select(CowTagID, NbSp:Vol8D, Season, Parameters, {{myparam}}) %>% 
    pivot_wider(names_from = Parameters, values_from = {{myparam}})
  
  ParamType <- as.character(myparam)
  
  p_df <- tibble(Parameter = as.character(),
                 ParamType = as.character(),
                 Dependent = as.character(),
                 Season = as.character(),
                 pvalue1 = as.numeric(),
                 pvalue2 = as.numeric(),
                 r_squared = as.numeric(),
                 adj_r_squared = as.numeric())
  
  for(i in 2:6){ # 2:6 are our dependent variables NbSp, NbFEs, Vol8D
  
    for(j in 8:ncol(Param_data)){
      
      Param_data <- Param_data %>% 
        filter(Season == myseason)
      Season <- myseason
      
      Parameter <- colnames(Param_data)[j]
      Dependent <- colnames(Param_data)[i] # select dependent parameter
      
      if(is.na(Param_data[1,j])){ # some variables have no data in the wet season, so skip these variables
        p_df <- p_df
      } else {
        mod <- lm(paste(Dependent, "~ poly(", Parameter, ",2)"), data = Param_data)
        pvalue1 <- summary(mod)[4]$coefficients[11]
        pvalue2 <- summary(mod)[4]$coefficients[12]
        r_squared <- summary(mod)[8]$r.squared
        adj_r_squared <- summary(mod)[9]$adj.r.squared

        temp <- as_tibble(cbind(Parameter, ParamType, Dependent, Season, 
                          pvalue1, pvalue2, r_squared, adj_r_squared)) %>%
              mutate(pvalue1 = as.numeric(pvalue1),
                     pvalue2 = as.numeric(pvalue2),
                     r_squared = as.numeric(r_squared),
                     adj_r_squared = as.numeric(adj_r_squared))

        p_df <- p_df %>%
          rbind(temp) 
      }
    }
  }
  return(p_df)
}

# pvalue plots
pvalplot <- function(mydata, season = "Dry"){

  mydata <- mydata %>% 
    pivot_longer(cols = c(pvalue1, pvalue2), names_to = "p1_p2", values_to = "pvalue")
  # pvalue1 is the same in Linear and Poly2
  
  season <- enquo(season)
  
  plot1 <- mydata %>% 
    filter(Dependent == "NbFEs" | Dependent == "NbSp" | Dependent == "Vol8D") %>% 
    filter(Season == {{season}}) %>% 
    ggplot(aes(x = Parameter, y  = pvalue)) +
    geom_col(aes(fill = p1_p2),
             position = "dodge") +
    theme_bw() +
    geom_hline(yintercept = 0.05, color = "black", size = 1) +
    theme(axis.text.x = element_text(angle = 90),
          legend.position = "top") +
    facet_wrap(~Dependent, scales = "free")
  
  return(plot1)
}


```

# INCLUDING VSEEP

## TURBINARIA NUTRIENT LOADING
```{r}
plotfun(y = NbSp, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = NbSp, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)

plotfun(y = NbFEs, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = NbFEs, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)

plotfun(y = Vol8D, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = Vol8D, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)


# get pvalues
pTO<-pvalpoly(mydata = myturb, myparam = "Values", myseason = "June2022")
pTO

pvalplot(mydata = pTO, season = "June2022")
```



## MAXIMUM VALUES
```{r}

plotfun(y = NbSp, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")


# get pvalues
Max3<-pvalpoly(myparam = "Maximum", myseason = "Dry")
Max4<-pvalpoly(myparam = "Maximum", myseason = "Wet")


Max <- full_join(Max3, Max4)
Max

pvalplot(mydata = Max, season = "Dry")
pvalplot(mydata = Max, season = "Wet")

```

## MINIMUM VALUES
```{r}

plotfun(y = NbSp, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
Min3<-pvalpoly(myparam = "Minimum", myseason = "Dry")
Min4<-pvalpoly(myparam = "Minimum", myseason = "Wet")

Min <- full_join(Min3, Min4)
Min



pvalplot(mydata = Min, season = "Dry")
pvalplot(mydata = Min, season = "Wet")
```


## RANGE VALUES
```{r}

plotfun(y = NbSp, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
Ran3<-pvalpoly(myparam = "Range", myseason = "Dry")
Ran4<-pvalpoly(myparam = "Range", myseason = "Wet")

Ran <- full_join(Ran3, Ran4) 
Ran


pvalplot(mydata = Ran, season = "Dry")
pvalplot(mydata = Ran, season = "Wet")
```
## MEAN VALUES BY SEASON
```{r}

plotfun(y = NbSp, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
MeanS3<-pvalpoly(myparam = "MeanSeasonal", myseason = "Dry")
MeanS4<-pvalpoly(myparam = "MeanSeasonal", myseason = "Wet")

MeanS <- full_join(MeanS3, MeanS4) 
MeanS


pvalplot(mydata = MeanS, season = "Dry")
pvalplot(mydata = MeanS, season = "Wet")
```

## MEAN VALUES (ALL SEASONS)
```{r}

plotfun(y = NbSp, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
MeanA3<-pvalpoly(myparam = "MeanAll", myseason = "Dry")
MeanA4<-pvalpoly(myparam = "MeanAll", myseason = "Wet")

MeanA <- full_join(MeanA3, MeanA4) 
MeanA

pvalplot(mydata = MeanA, season = "Dry")
pvalplot(mydata = MeanA, season = "Wet")
```



## CV VALUES BY SEASON
```{r}

plotfun(y = NbSp, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
CVS3<-pvalpoly(myparam = "CVSeasonal", myseason = "Dry")
CVS4<-pvalpoly(myparam = "CVSeasonal", myseason = "Wet")

CVS <- full_join(CVS3, CVS4)
CVS

pvalplot(mydata = CVS, season = "Dry")
pvalplot(mydata = CVS, season = "Wet")
```


## CV VALUES (ALL SEASONS)
```{r}

plotfun(y = NbSp, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
CVA3<-pvalpoly(myparam = "CVAll", myseason = "Dry")
CVA4<-pvalpoly(myparam = "CVAll", myseason = "Wet")

CVA <- full_join(CVA3, CVA4)
CVA

pvalplot(mydata = CVA, season = "Dry")
pvalplot(mydata = CVA, season = "Wet")
```



# NOT INCLUDING VSEEP


## Remove VSEEP from Full_data 
```{r}
Full_data <- Full_data %>% 
  filter(CowTagID!= "VSEEP")
myturb <- myturb %>% 
  filter(CowTagID != "VSEEP")
```


## Function: Visualize linear and nonlinear regressions: Site Characteristics
```{r}
paramPlot <- function(mydata = Full_data, ParamType, PTname, myformula = "y~x"){
  
  # isolate the desired paramtype (ex. Max, Min, Mean)
  data <- mydata %>% 
    select(-c(NbSp:Vol8D,lat, lon))
  
  myparamdata <- data %>% 
    select(CowTagID, Season, Parameters, Maximum)# {{ParamType}})
  
  data <- data[,1:13]
  data <- full_join(data, myparamdata)
  
  data <- data %>%
    pivot_wider(names_from = Parameters, values_from = Maximum) %>% # {{ParamType}}) %>% 
    relocate(Season, .after = CowTagID)
  
  p <- list() # empty list for saving plots
  
  for(i in 3:ncol(data)){
  # isolate dependent variable
  myDep <- data %>% 
    select(CowTagID, Season) %>%  # will join back after pivoting remaining variables
    cbind(data[,i]) %>% 
    drop_na()
  
  newdata <- data %>% 
    select(-c(colnames(data[,i]))) %>% # remove dependent var to not pivot on itself
    pivot_longer(cols = 3:ncol(.), names_to = "Parameters", values_to = "Values") %>% 
    right_join(myDep)  # bring back dependent var
  
  yplot <- as.name(colnames(newdata[,5])) 
  
  p[[i]] <- newdata %>% 
    ggplot(aes(x = Values, y = !!yplot, color = Season)) +
    geom_point() +
    geom_smooth(method = "lm", formula = myformula) +
    theme_bw() +
    theme(legend.position = "top",
          legend.direction = "horizontal") +
    facet_wrap(~Parameters, scales = "free")
  }

  # Save all plots in a single pdf
  pdf(here("Output",paste0("V_site_regressions_", PTname ,".pdf")), onefile = TRUE)
  for (i in 1:length(p)) {
    tplot <- p[[i]]
    print(tplot)
  }
  dev.off()

return(p)
}
```




## TURBINARIA NUTRIENT LOADING
```{r}
plotfun(y = NbSp, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = NbSp, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)

plotfun(y = NbFEs, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = NbFEs, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)

plotfun(y = Vol8D, x = Values, myfacet = Parameters, myformula = "y~x", mydata = myturb)
plotfun(y = Vol8D, x = Values, myfacet = Parameters, myformula = "y~poly(x,2)", mydata = myturb)


# get pvalues
pTO<-pvalpoly(mydata = myturb, myparam = "Values", myseason = "June2022")
pTO

pvalplot(mydata = pTO, season = "June2022")
```



## MAXIMUM VALUES
```{r}

plotfun(y = NbSp, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Maximum, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Maximum, myfacet = Parameters, myformula = "y~poly(x,2)")


# get pvalues
Max3<-pvalpoly(myparam = "Maximum", myseason = "Dry")
Max4<-pvalpoly(myparam = "Maximum", myseason = "Wet")


Max <- full_join(Max3, Max4)
Max

pvalplot(mydata = Max, season = "Dry")
pvalplot(mydata = Max, season = "Wet")

```

```{r}
p1 <- paramPlot(mydata = Full_data, ParamType = Maximum, PTname = "Maximum")
p1b <- paramPlot(mydata = Full_data, ParamType = Maximum, PTname = "Maximum_poly", myformula = "y~poly(x,2)")
```


## MINIMUM VALUES
```{r}

plotfun(y = NbSp, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Minimum, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Minimum, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
Min3<-pvalpoly(myparam = "Minimum", myseason = "Dry")
Min4<-pvalpoly(myparam = "Minimum", myseason = "Wet")

Min <- full_join(Min3, Min4)
Min



pvalplot(mydata = Min, season = "Dry")
pvalplot(mydata = Min, season = "Wet")
```

```{r}
p2 <- paramPlot(mydata = Full_data, ParamType = Minimum, PTname = "Minimum")
p2b <- paramPlot(mydata = Full_data, ParamType = Minimum, PTname = "Minimum_poly", myformula = "y~poly(x,2)")
```

## RANGE VALUES
```{r}

plotfun(y = NbSp, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = Range, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = Range, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
Ran3<-pvalpoly(myparam = "Range", myseason = "Dry")
Ran4<-pvalpoly(myparam = "Range", myseason = "Wet")

Ran <- full_join(Ran3, Ran4) 
Ran


pvalplot(mydata = Ran, season = "Dry")
pvalplot(mydata = Ran, season = "Wet")
```

```{r}
p3 <- paramPlot(mydata = Full_data, ParamType = Range, PTname = "Range")
p3b <- paramPlot(mydata = Full_data, ParamType = Range, PTname = "Range_poly", myformula = "y~poly(x,2)")
```

## MEAN VALUES BY SEASON
```{r}

plotfun(y = NbSp, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = MeanSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = MeanSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
MeanS3<-pvalpoly(myparam = "MeanSeasonal", myseason = "Dry")
MeanS4<-pvalpoly(myparam = "MeanSeasonal", myseason = "Wet")

MeanS <- full_join(MeanS3, MeanS4) 
MeanS


pvalplot(mydata = MeanS, season = "Dry")
pvalplot(mydata = MeanS, season = "Wet")
```

```{r}
p4 <- paramPlot(mydata = Full_data, ParamType = MeanSeasonal, PTname = "MeanSeasonal")
p4b <- paramPlot(mydata = Full_data, ParamType = MeanSeasonal, PTname = "MeanSeasonal_poly", myformula = "y~poly(x,2)")
```

## MEAN VALUES (ALL SEASONS)
```{r}

plotfun(y = NbSp, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = MeanAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = MeanAll, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
MeanA3<-pvalpoly(myparam = "MeanAll", myseason = "Dry")
MeanA4<-pvalpoly(myparam = "MeanAll", myseason = "Wet")

MeanA <- full_join(MeanA3, MeanA4) 
MeanA

pvalplot(mydata = MeanA, season = "Dry")
pvalplot(mydata = MeanA, season = "Wet")
```

```{r}
p5 <- paramPlot(mydata = Full_data, ParamType = MeanAll, PTname = "MeanAll")
p5b <- paramPlot(mydata = Full_data, ParamType = MeanAll, PTname = "MeanAll_poly", myformula = "y~poly(x,2)")
```

## CV VALUES BY SEASON
```{r}

plotfun(y = NbSp, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = CVSeasonal, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = CVSeasonal, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
CVS3<-pvalpoly(myparam = "CVSeasonal", myseason = "Dry")
CVS4<-pvalpoly(myparam = "CVSeasonal", myseason = "Wet")

CVS <- full_join(CVS3, CVS4)
CVS

pvalplot(mydata = CVS, season = "Dry")
pvalplot(mydata = CVS, season = "Wet")
```

```{r}
p6 <- paramPlot(mydata = Full_data, ParamType = CVSeasonal, PTname = "CVSeasonal")
p6b <- paramPlot(mydata = Full_data, ParamType = CVSeasonal, PTname = "CVSeasonal_poly", myformula = "y~poly(x,2)")
```

## CV VALUES (ALL SEASONS)
```{r}

plotfun(y = NbSp, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbSp, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = NbFEs, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = NbFEs, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

plotfun(y = Vol8D, x = CVAll, myfacet = Parameters, myformula = "y~x")
plotfun(y = Vol8D, x = CVAll, myfacet = Parameters, myformula = "y~poly(x,2)")

# get pvalues
CVA3<-pvalpoly(myparam = "CVAll", myseason = "Dry")
CVA4<-pvalpoly(myparam = "CVAll", myseason = "Wet")

CVA <- full_join(CVA3, CVA4)
CVA

pvalplot(mydata = CVA, season = "Dry")
pvalplot(mydata = CVA, season = "Wet")
```

```{r}
p7 <- paramPlot(mydata = Full_data, ParamType = CVAll, PTname = "CVAll")
p7b <- paramPlot(mydata = Full_data, ParamType = CVAll, PTname = "CVAll_poly", myformula = "y~poly(x,2)")
```



