---
title: "Community vs SGD Regression"
author: "Danielle Barnas"
date: "2022-10-14"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = TRUE)
```

# Install necessary libraries
```{r, eval = F}
if("tidyverse" %in% rownames(installed.packages()) == FALSE){install.packages("tidyverse")}
if("here" %in% rownames(installed.packages()) == FALSE){install.packages("here")}
if("ggrepel" %in% rownames(installed.packages()) == FALSE){install.packages("ggrepel")}
if("PNWColors" %in% rownames(installed.packages()) == FALSE){install.packages("PNWColors")}
if("vegan" %in% rownames(installed.packages()) == FALSE){install.packages("vegan")}
if("pairwiseAdonis" %in% rownames(installed.packages()) == FALSE){ devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")}
if("patchwork" %in% rownames(installed.packages()) == FALSE){install.packages("patchwork")}
if("ggmap" %in% rownames(installed.packages()) == FALSE){install.packages("ggmap")}
```


# Load Libraries
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(PNWColors)
library(vegan)
library(pairwiseAdonis)
library(patchwork)
library(ggmap)

set.seed(7)
```


# Read in data
```{r, message = F}
#div <- read_csv(here("Data","Surveys","Species_Diversity.csv")) # cannot calculate species div with % cover data
rich <- read_csv(here("Data","Surveys","Species_Richness.csv"))
meta <- read_csv(here("Data", "Full_Metadata.csv"))
comp <- read_csv(here("Data","Surveys","Species_Composition_2022.csv"))
taxa <- read_csv(here("Data","Surveys","Distinct_Taxa.csv"))
chem <- read_csv(here("Data","Biogeochem", "Nutrient_Processed_CV.csv"))
means <- read_csv(here("Data","Biogeochem", "Nutrient_Processed_meanSeason.csv"))
ctTemp <- read_csv(here("Data","Biogeochem", "CT_files", "cleaned_CT.csv"))

# create color palette for plotting
mypalette <- pnw_palette(name="Bay")
midpalette <- pnw_palette(name = "Bay", n = 30)
largepalette <- pnw_palette(name = "Bay", n = 100)

## create palette
pretty_palette <- c("#d8a7b1", "#e9ddd4","#b9b7bd", "#67595e", "#ef7c8e", "#fae8e0", "#747474") # 1 is rosewater, 2 is dusty rose, 3 is pewter, 4 is coffee pot, 5 is hot pink, 6 is cream, 7 is light grey

```


# Process and Join data
```{r}
# reduce metadata
meta <- meta %>% 
  drop_na(lat) %>% 
  select(Location, CowTagID, LiveCoral:Sand, dist_to_seep_m, adj_CT_depth_cm, meanRugosity) %>% 
  filter(Location == "Varari",
         CowTagID != "VSEEP",
         CowTagID != "V13")

cvtemp <- ctTemp %>% 
  select(CowTagID, CVTemp)
meantemp <- ctTemp %>% 
  select(CowTagID, meanTemp)

# get means across seasons
means <- means %>% 
  group_by(Location, CowTagID) %>% 
  summarise_at(vars(lat:N_percent), .funs = mean, na.rm =T) %>% 
  left_join(rich) %>% 
  left_join(meta) %>% 
  relocate(spRichness, .after = CowTagID) %>%
  filter(Location == "Varari",
         CowTagID != "VSEEP",
         CowTagID != "V13") %>% 
  ungroup() %>% 
  left_join(meantemp) %>% 
  select(-Temperature) %>% 
  relocate(meanTemp, .after = Salinity) %>% 
  rename(Temperature = meanTemp)

# join dfs
Full_data <- rich %>% 
  full_join(chem) %>% 
  drop_na(lat) %>% # remove largely incomplete datasets
  filter(CowTagID != "C14") %>%  # not surveyed
  left_join(meta) %>% 
  filter(Location == "Varari",
         CowTagID != "VSEEP",
         CowTagID != "V13") %>% 
  left_join(cvtemp) %>% 
  select(-Temperature) %>% 
  relocate(CVTemp, .after = Salinity) %>% 
  rename(Temperature = CVTemp)


comp <- comp %>% 
  select(Location, CowTagID, Taxa, SpeciesCounts) %>% 
  mutate(Taxa = if_else(Taxa == 'Bare Rock', "Hard Substrate", Taxa),
         Taxa = if_else(Taxa == 'Rubble', "Hard Substrate", Taxa)) %>% 
  group_by(Location, CowTagID, Taxa) %>% 
  mutate(SpeciesCounts = sum(SpeciesCounts)) %>% 
  distinct() %>% 
  ungroup() %>% # ungroup by taxa to regroup by site
  group_by(Location, CowTagID) %>% 
  mutate(totalcount = sum(SpeciesCounts),
         sp_percent = SpeciesCounts / totalcount*100) %>% 
  select(-totalcount) %>% 
  left_join(Full_data) %>% 
  ungroup() %>% 
  filter(Location == "Varari", # only Varari for now
         CowTagID != "V13", CowTagID != "VSEEP") # removed from biogeochem analyses

# rearrange
Full_data <- Full_data %>% 
  relocate(Location, .before = CowTagID)

# only Varari for now
# remove seep and V13 point upstream for further analyses
Full_data <- Full_data %>% 
  filter(Location == "Varari") %>% 
  filter(CowTagID != "VSEEP",
         CowTagID != "V13")

# pivot longer by categories
longsub<- Full_data %>% 
  select(-c(lat, lon)) %>%  
  pivot_longer(cols = c(LiveCoral:Sand), names_to = "Substrate", values_to = "subVal")

```


# Functions

## Create function to order cowtagID's when plotting
```{r}
order <- function(data = Full_data, param, by){

  param<- enquo(param)
  by<- enquo(by)
  
  orderParam <- data %>%
     arrange(!!param) %>%
     distinct(!!by) %>% 
     mutate(CowTagID = as_factor(as.character(!!by)))

  # sort factor levels by N_percent
  ParamLevels <- paste(orderParam$CowTagID)
  # assign order to factor levels by % N
  data$CowTagID <- factor(data$CowTagID, levels = ParamLevels)
  # levels(data$CowTagID) # check

return(data)
}

## example: distance to seep
order(param = dist_to_seep_m, by = CowTagID)$CowTagID
Full_data$CowTagID <- order(param = dist_to_seep_m, by = CowTagID)$CowTagID

```



### Create function for geom_point plots with facet_wrapping
```{r}

# create ggplot function
groupplot <- function(mydata = Full_data, x, y, fw) {

  x<-enquo(x)
  y<-enquo(y)

  plot <- ggplot(data = mydata,
                 aes(y = !!y,
                     x = !!x,
                     color = !!enquo(fw))) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(enquo(fw), scales = "free_y") +
  labs(y = paste(as_label(y)),
       x = paste(as_label(x))) +
  theme_bw()+
  theme(legend.direction = "horizontal",
        legend.position = "top")
  
  return(plot)
}
```

### Create function for geom_col stacked
```{r}

# create ggplot function
stackplot <- function(mydata, x = CowTagID, y, colfill) {

  x<-enquo(x)
  y<-enquo(y)

  plot <- ggplot(data = mydata,
                 aes(y = !!y,
                     x = !!x,
                     fill = !!enquo(colfill))) +
  geom_col(position = "stack") +
  labs(y = paste(as_label(y)),
       x = paste(as_label(x))) +
  theme_bw()+
  theme(legend.position = "right")
  
  return(plot)
}
```

### Create function for geom_col facet_wrapped
```{r}

# create ggplot function
colplot <- function(mydata, x = CowTagID, y) {

  x<-enquo(x)
  y<-enquo(y)

  plot <- ggplot(data = mydata,
                 aes(y = !!y,
                     x = !!x)) +
  geom_col() +
  labs(y = paste(as_label(y)),
       x = paste(as_label(x))) +
  theme_bw()+
  theme(legend.position = "none")
  
  return(plot)
}
```


# Distance to seep regressions

## Plotting 
### All biogeochem ~ distance to seep
```{r}
longChem <- Full_data %>% 
  pivot_longer(cols = c(Salinity:N_percent), values_to = "Values", names_to = "Parameter") 
longChem %>% 
  drop_na(Values, dist_to_seep_m) %>%
  ggplot(aes(y = Values,
             x = dist_to_seep_m)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Parameter, scales = "free_y") +
  theme_bw()+
  theme(legend.direction = "none") +
  scale_fill_manual(values = largepalette)


```


### Species richness ~ distance to seep
```{r}
## DISTANCE TO SEEP

groupplot(mydata = Full_data, x = dist_to_seep_m, y = spRichness, fw = NA) +
    geom_text_repel(aes(label = CowTagID), size = 3)+
  scale_color_manual(values = mypalette)

anova(lm(data = Full_data, spRichness ~ dist_to_seep_m)) # significant **
```

### Stacked % Cover substrate ~ % N
```{r}

# order % cover by N_percent
longsub$CowTagID <- order(data = longsub, param = dist_to_seep_m, by = CowTagID)$CowTagID

# ggplot
stackplot(mydata = longsub, x = CowTagID, y = subVal, colfill = Substrate) +
  labs(subtitle = "Distance from seep, Near to Far") +
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = pretty_palette[c(1, 3, 6, 4)])
  
```


### Stacked % Cover species ~ % N
```{r}

# order % cover by N_percent
comp$CowTagID <- order(data = comp, param = VisibleHumidic_Like, by = CowTagID)$CowTagID

# ggplot
stackplot(mydata = comp %>% filter(Taxa != "Hard Substrate", Taxa != "Sand"), x = CowTagID, y = sp_percent, colfill = Taxa) +
  labs(subtitle = "Low to High Visible Humidics-like") +
  scale_fill_manual(values = largepalette) +
  theme(legend.text = element_text(size = 2),
        axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = largepalette)
```


### Stacked % Cover species ~ distance to seep
```{r}

# order % cover by N_percent
comp$CowTagID <- order(data = comp, param = dist_to_seep_m, by = CowTagID)$CowTagID

# ggplot
stackplot(mydata = comp %>% filter(Taxa != "Hard Substrate", Taxa != "Sand", CowTagID != "V13"), x = CowTagID, y = sp_percent, colfill = Taxa) +
  labs(subtitle = "Distance from seep (m)") +
  scale_fill_manual(values = largepalette) +
  theme(legend.text = element_text(size = 2),
        axis.text.x = element_text(angle = 90))
```


# PCA: CV of biogeochemical characteristics

## PCA
```{r}
# extract parameters for the PCA, excluding CT depth
pca_fulldata <- Full_data %>% 
  select(Salinity, Temperature, pH, 
         Phosphate_umolL, Silicate_umolL, NN_umolL, Ammonia_umolL,
         M_C, HIX, VisibleHumidic_Like, Tryptophan_Like, Tyrosine_Like
         #del15N, C_N, N_percent,  remove because of mean values
         #meanRugosity
         ) # removed TA, substrate, distance to seep, and CT depth, and rugosity

# Run the PCA
pca_full <- prcomp(pca_fulldata, scale. = TRUE, center = TRUE)

# Proportion of variance
summary(pca_full)
# PC1: 24.97 %
# PC2: 19.68 %

# Extract the scores and loadings
PC_scores_full <-as_tibble(pca_full$x[,1:2]) # scores for each sample location
PC_loadings_full<-as_tibble(pca_full$rotation) %>% # loadings for each parameter
  bind_cols(labels = rownames(pca_full$rotation))

# Put it with all the original data
pca_fulldata_all<-Full_data %>% 
  bind_cols(PC_scores_full)

```

## scores plot
```{r}
p1_full <- pca_fulldata_all %>%
  ggplot(aes(x = PC1, 
             y = PC2))+
  geom_point(shape = 21,
             color = "white",
             fill = "white",
             size = 3) +
  geom_text_repel(aes(label = CowTagID)) +
  scale_colour_hue(l = 45)+ # l: luminance/lightness 0-100
  scale_fill_hue(l = 45)+ 
  theme_bw()+
  theme(legend.position = "none", 
        panel.background = element_rect(fill = pretty_palette[3]),  
        panel.grid = element_blank()) +
  labs(title = "Varari Scores Plot (CV Values)",
       x = "PC1 - 25.0%", y = "PC2 - 19.7%")

p1_full
```


## loadings plot
```{r}
# loadings plot 
p1_full_loading <- PC_loadings_full %>%
  ggplot(aes(x = PC1, y = PC2, label = labels)) + # labels are the parameters
    geom_segment(aes(x = 0,
                     y = 0,
                     xend = PC1*10,
                     yend = PC2*10),
      arrow = arrow(length = unit(0.2,"cm"),), 
      color = pretty_palette[1],
      size = 1) +
  annotate("text", 
           x = PC_loadings_full$PC1*10+0.5, 
           y = PC_loadings_full$PC2*10+0.5,
           label = PC_loadings_full$labels,
           size = 3) +
  #coord_cartesian(xlim = c(-4, 5), ylim = c(-6, 5)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(title = "Varari Loadings (CV Values)",
       x = "PC1 - 25.0%", y = "PC2 - 19.7%")

p1_full_loading
ggsave(here("Output", "PCA", "CV_PCAloadings.png"), p1_full_loading, height = 14, width = 10)

p1_full_loadingBlank <- PC_loadings_full %>%
  ggplot(aes(x = PC1, y = PC2, label = labels)) + # labels are the parameters
    geom_segment(aes(x = 0,
                     y = 0,
                     xend = PC1*10,
                     yend = PC2*10),
      arrow = arrow(length = unit(0.2,"cm"),), 
      color = pretty_palette[1],
      size = 1) +
  #coord_cartesian(xlim = c(-4, 5), ylim = c(-6, 5)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(title = "Varari Loadings (CV Values)",
       x = "PC1 - 25.0%", y = "PC2 - 19.7%")

p1_full_loadingBlank
ggsave(here("Output", "PCA", "CV_PCAloadings_blank.png"), p1_full_loadingBlank, height = 14, width = 10)
```


## Save plots together
```{r}
### Patch PCA plots
VarariPCA <- p1_full + p1_full_loading + 
  patchwork::plot_annotation("Varari PCA and Loadings (CV Values)", 
                             theme = theme(plot.title = element_text(size = rel(1.5), 
                                                                     face = "bold", 
                                                                     hjust = 0.5, 
                                                                     margin = margin(t = 10, b = 20,
                                                                                     unit = "pt"))))

ggsave(plot = VarariPCA, filename = here("Output","PCA","CV_PCA.png"), width = 14, height = 10)
VarariPCA
```

### species richness ~ PC scores
```{r}
#pca_fulldata_all

anova(lm(data = pca_fulldata_all, spRichness ~ PC1))
anova(lm(data = pca_fulldata_all, spRichness ~ PC2))

pca_fulldata_all %>% 
  groupplot(x = PC1, y = spRichness, fw = Location) +
  scale_color_manual(values = mypalette) +
  theme(legend.position = "none")

pca_fulldata_all %>% 
  groupplot(x = PC2, y = spRichness, fw = Location) +
  scale_color_manual(values = mypalette) +
  theme(legend.position = "none") +
  geom_text_repel(aes(label = CowTagID))


summary(lm(data = pca_fulldata_all, PC1 ~ dist_to_seep_m))
summary(lm(data = pca_fulldata_all, PC2 ~ dist_to_seep_m))

```


### Graph survey locations colored by PC1

```{r}

# mean lat and long for the maps
LocationGPS <- pca_fulldata_all %>%
  group_by(Location) %>% # varari vs cabral
  summarise(lon = median(lon, na.rm = TRUE) + 0.00007,
            lat = median(lat, na.rm = TRUE))


# Varari
VarariBaseMap<-get_map(LocationGPS %>%
                         select(lon,lat), maptype = 'satellite', zoom = 19)

# base map
# Varari
VmapSites <- ggmap(VarariBaseMap) +
  geom_point(data = pca_fulldata_all,
             aes(x = lon, y = lat),
             color = "white",
             size = 2) + 
  labs(x = "Longitude", y = "Latitude",  #label x and y axes
        title = "Varari Sample Locations") +
  geom_label(data = pca_fulldata_all,
             aes(x = lon, y = lat,
                 label = CowTagID),
             size = 2, alpha = 0.8)

VmapSites

### Map Varari colored by PC1

# isolate Varari seep
VSeepPt <- chem %>% filter(CowTagID == "VSEEP") %>% select(CowTagID, lat, lon)
  

reg_varari_map <- ggmap(VarariBaseMap) +
  geom_point(data = pca_fulldata_all,
             aes(x = lon, y = lat, fill = PC1),
             shape = 21,
             color = "black",
             size = 3) + # set point aesthetics
  scale_fill_gradient(low = "white", high = pretty_palette[5]) + # set color scale
  geom_point(data = VSeepPt, # star the seepage point
             aes(x = lon, y = lat),
             shape = 18, # selects the palette color corresponding to the associated cluster number
             color = "white",
             size = 4) +
  geom_label(data = VSeepPt,
             aes(x = lon, y = lat,
                 label = CowTagID),
             size = 3) + 
             #nudge_y = 0.00012) +
  labs(x = "Longitude", y = "Latitude",  #label x and y axes
        title = "Varari Survey Locations, CV Values",
       fill = "PC1 (25.0%)")

reg_varari_map
ggsave(here("Output", "PCA", "CV_map_colorPC1.png"), reg_varari_map, height = 10, width = 10)

```



# PCA: Mean of biogeochemical characteristics and TO

## PCA
```{r}
# extract parameters for the PCA, excluding CT depth
pca_fulldata2 <- means %>% 
  select(Salinity, Temperature, pH, 
         Phosphate_umolL, Silicate_umolL, NN_umolL, Ammonia_umolL,
         M_C, HIX, VisibleHumidic_Like, Tryptophan_Like, Tyrosine_Like,
         del15N, C_N, N_percent) # removed TA, substrate, distance to seep, and CT depth and rugosity

# Run the PCA
pca_full2 <- prcomp(pca_fulldata2, scale. = TRUE, center = TRUE)

# Proportion of variance
summary(pca_full2)
# PC1: 37.82 %
# PC2: 16.48 %

# Extract the scores and loadings
PC_scores_full2 <-as_tibble(pca_full2$x[,1:2]) # scores for each sample location
PC_loadings_full2 <- as_tibble(pca_full2$rotation) %>% # loadings for each parameter
  bind_cols(labels = rownames(pca_full2$rotation))

# Put it with all the original data
pca_fulldata_all2<-means %>% 
  bind_cols(PC_scores_full2)

```

## scores plot
```{r}
p2_full<-pca_fulldata_all2 %>%
  ggplot(aes(x = PC1, 
             y = PC2))+
  geom_point(shape = 21,
             color = "white",
             fill = "white",
             size = 3) +
  geom_text_repel(aes(label = CowTagID)) +
  scale_colour_hue(l = 45)+ # l: luminance/lightness 0-100
  scale_fill_hue(l = 45)+ 
  theme_bw()+
  theme(legend.position = "none", 
        panel.background = element_rect(fill = pretty_palette[3]),
        panel.grid = element_blank()) +
  labs(title = "Varari Scores Plot (Mean Values)",
       x = "PC1 = 37.6%", y = "PC2 - 16.5%")

p2_full
```


## loadings plot
```{r}
# loadings plot 
p2_full_loading <- PC_loadings_full2 %>%
  ggplot(aes(x = PC1, y = PC2, label = labels)) + # labels are the parameters
    geom_segment(aes(x = 0,
                     y = 0,
                     xend = PC1*10,
                     yend = PC2*10),
      arrow = arrow(length = unit(0.2,"cm"),), 
      color = pretty_palette[1],
      size = 1) +
  annotate("text", 
           x = PC_loadings_full2$PC1*10.5+0.3, 
           y = PC_loadings_full2$PC2*10.5+0.5,
           label = PC_loadings_full2$labels,
           size = 3) +
  #coord_cartesian(xlim = c(-4, 5), ylim = c(-5, 5)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(title = "Varari Loadings (Mean Values)",
       x = "PC1 = 37.6%", y = "PC2 - 16.5%")

p2_full_loading
ggsave(here("Output", "PCA", "Means_PCAloadings.png"), p2_full_loading, height = 14, width = 10)

p2_full_loadingBlank <- PC_loadings_full2 %>%
  ggplot(aes(x = PC1, y = PC2, label = labels)) + # labels are the parameters
    geom_segment(aes(x = 0,
                     y = 0,
                     xend = PC1*10,
                     yend = PC2*10),
      arrow = arrow(length = unit(0.2,"cm"),), 
      color = pretty_palette[1],
      size = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(title = "Varari Loadings (Mean Values)",
       x = "PC1 = 37.6%", y = "PC2 - 16.5%")

p2_full_loadingBlank
ggsave(here("Output", "PCA", "Means_PCAloadings_blank.png"), p2_full_loadingBlank, height = 14, width = 10)

```


## Save plots together
```{r}
### Patch PCA plots
VarariPCA2 <- p2_full + p2_full_loading + 
  patchwork::plot_annotation("Varari PCA and Loadings (Mean Values)", 
                             theme = theme(plot.title = element_text(size = rel(1.5), 
                                                                     face = "bold", 
                                                                     hjust = 0.5, 
                                                                     margin = margin(t = 10, b = 20,
                                                                                     unit = "pt"))))

ggsave(plot = VarariPCA2, filename = here("Output","PCA","Means_PCA.png"), width = 14, height = 10)
VarariPCA2
```


### species richness ~ PC scores
```{r}
#pca_fulldata_all

anova(lm(data = pca_fulldata_all2, spRichness ~ PC1))
anova(lm(data = pca_fulldata_all2, spRichness ~ PC2))

pca_fulldata_all2 %>% 
  groupplot(x = PC1, y = spRichness, fw = NA) +
  scale_color_manual(values = mypalette)

pca_fulldata_all2 %>% 
  groupplot(x = PC2, y = spRichness, fw = NA) +
  scale_color_manual(values = mypalette)


summary(lm(data = pca_fulldata_all2, PC1 ~ dist_to_seep_m))
summary(lm(data = pca_fulldata_all2, PC2 ~ dist_to_seep_m))

```

### Graph survey locations colored by PC1

PC1 Axis

```{r}

### Map Varari colored by PC1

reg_varari_map_means <- ggmap(VarariBaseMap) +
  geom_point(data = pca_fulldata_all2,
             aes(x = lon, y = lat, fill = PC1),
             shape = 21,
             color = "black",
             size = 3) + # set point aesthetics
  scale_fill_gradient(low = "white", high = pretty_palette[5]) + # set color scale
  geom_point(data = VSeepPt, # star the seepage point
             aes(x = lon, y = lat),
             shape = 18, # selects the palette color corresponding to the associated cluster number
             color = "white",
             size = 4) +
  geom_label(data = VSeepPt,
             aes(x = lon, y = lat,
                 label = CowTagID),
             size = 3) + 
             #nudge_y = 0.00012) +
  labs(x = "Longitude", y = "Latitude",  #label x and y axes
        title = "Varari Survey Locations, Mean Values",
       fill = "PC1 (37.6%)")

reg_varari_map_means
ggsave(here("Output", "PCA", "means_map_colorPC1.png"), reg_varari_map_means, height = 10, width = 10)
```

PC2 Axis

```{r}
reg_varari_map_means2 <- ggmap(VarariBaseMap) +
  geom_point(data = pca_fulldata_all2,
             aes(x = lon, y = lat, fill = PC2),
             shape = 21,
             color = "black",
             size = 3) + # set point aesthetics
  scale_fill_gradient(low = "white", high = pretty_palette[5]) + # set color scale
  geom_point(data = VSeepPt, # star the seepage point
             aes(x = lon, y = lat),
             shape = 18, # selects the palette color corresponding to the associated cluster number
             color = "white",
             size = 4) +
  geom_label(data = VSeepPt,
             aes(x = lon, y = lat,
                 label = CowTagID),
             size = 3) + 
             #nudge_y = 0.00012) +
  labs(x = "Longitude", y = "Latitude",  #label x and y axes
        title = "Varari Survey Locations, Mean Values",
       fill = "PC2 (16.5%)")

reg_varari_map_means2
ggsave(here("Output", "PCA", "means_map_colorPC2.png"), reg_varari_map_means2, height = 10, width = 10)

```



# Normalize to rugosity (and thereby substrate)

## Richness ~ Rugosity
```{r}
## plot
groupplot(mydata = Full_data, x = meanRugosity, y = spRichness, fw = Location) +
  geom_text_repel(aes(label = CowTagID)) +
  scale_color_manual(values = mypalette) +
  theme(legend.position = "none")

# A residual is the difference between an observed value and a predicted value in a regression model. Residual = Observed value – Predicted value
# An observation has a positive residual if its value is greater than the predicted value made by the regression line and a negative residual if its value is less than the predicted value made by the regression line.

#fit model
resMod <- lm(spRichness ~ meanRugosity, data=Full_data)

#view model summary
summary(resMod) 

#calculate the standardized residuals
res <- residuals(resMod)


#column bind standardized residuals back to original data frame
res_data <- Full_data %>% 
  cbind(res) %>% 
  relocate(res, .after = lon)

#column bind standardized residuals back to means data frame
res_data_means <- means %>% 
  cbind(res) %>% 
  relocate(res, .after = lon)

#plot predictor variable vs. standardized residuals
res_data %>% 
  ggplot(aes(x = meanRugosity,
             y = res)) +
  geom_point() +
  geom_text_repel(aes(label = CowTagID)) +
  theme_bw()

res_data %>% 
  ggplot(aes(x = Sand,
             y = res)) +
  geom_point() +
  geom_text_repel(aes(label = CowTagID)) +
  theme_bw()

# removed interaction with substrate as well
anova(lm(data = res_data, Sand ~ res)) 
anova(lm(data = res_data, DeadCoral ~ res))
anova(lm(data = res_data, LiveCoral ~ res))
anova(lm(data = res_data, Rubble ~ res))
```

```{r}
# plot residuals against richness
respval1 <- anova(lm(data = res_data, spRichness ~ res))[5]$'Pr(>F)'[1]

res_data %>% 
  ggplot(aes(x = res,
             y = spRichness)) +
  geom_point() +
  geom_text_repel(aes(label = CowTagID)) +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() + theme(panel.grid = element_blank()) +
  labs(y = "Residuals (Species Richness ~ Mean Rugosity)",
       x = "Species Richness",
       caption = paste("p-value < 8.37e-06")) #round(respval1,3)))


# plot residuals against distance parameter
respval2 <- anova(lm(data = res_data, res ~ dist_to_seep_m))[5]$'Pr(>F)'[1]

res_data %>% 
  ggplot(aes(x = dist_to_seep_m,
             y = res)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() + theme(panel.grid = element_blank()) +
  labs(y = "Residuals (Species Richness ~ Mean Rugosity)",
       x = "Distance to seepage point (m)",
       caption = paste("p-value:",round(respval2,3)))

```


```{r}
# plot residuals against richness
respval1 <- anova(lm(data = res_data, spRichness ~ res))[5]$'Pr(>F)'[1]
respval1

res_data %>% 
  ggplot(aes(x = res,
             y = spRichness)) +
  geom_point() +
  geom_text_repel(aes(label = CowTagID)) +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() + theme(panel.grid = element_blank()) +
  labs(y = "Residuals (Species Richness ~ Mean Rugosity)",
       x = "Species Richness",
       caption = paste("p-value =", respval1)) #round(respval1,3)))


# plot residuals against distance parameter
respval2 <- anova(lm(data = res_data, res ~ dist_to_seep_m))[5]$'Pr(>F)'[1]
respval2

res_data %>% 
  ggplot(aes(x = dist_to_seep_m,
             y = res)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() + theme(panel.grid = element_blank()) +
  labs(y = "Residuals (Species Richness ~ Mean Rugosity)",
       x = "Distance to seepage point (m)",
       caption = paste("p-value:",round(respval2,3)))

```



## Calculate p values for regressions

### CV Values wtih scaled parameters
```{r}
Param_data <- res_data %>% 
  relocate(spRichness, .after = res)

scaleParam <- as_tibble(scale(Param_data[,6:29], scale = TRUE, center = TRUE)) %>% 
  bind_cols(res_data %>% select(res)) %>% 
  relocate(res, .before = spRichness)



Residuals_pval <- tibble(Parameter = as.character(),
                        pvalue = as.numeric(),
                        adj_r_squared = as.numeric(),
                        beta = as.numeric())

for(i in 2:ncol(scaleParam)){
  
  Parameter <- colnames(scaleParam)[i] # select dependent parameter
  
  mod <- lm(paste(Parameter, "~", "res"), data = scaleParam)
  pvalue <- summary(mod)[4]$coefficients[8]
  r_squared <- summary(mod)[8]$r.squared
  beta <- summary(mod)$coefficients[2]
  
  temp <- as_tibble(cbind(Parameter, pvalue, r_squared)) %>% 
    mutate(pvalue = as.numeric(pvalue),
           r_squared = as.numeric(r_squared),
           beta = as.numeric(beta))
  
  Residuals_pval <- Residuals_pval %>% 
    rbind(temp)
}

# view table of pvalues
Residuals_pval

# view bar graph of pvalues 
Residuals_pval %>% 
  arrange(beta) %>% 
  ggplot(aes(x = fct_reorder(Parameter, beta), y = beta)) +
  geom_col() +
  geom_hline(yintercept = 0.0) +
  theme_bw() +
  labs(y = "effect size: ~ Residuals from lm(Richness ~ Rugosity)",
       title = "CV of biogeochemical values across Seasons") +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
res_data %>% 
  ggplot(aes(x = res,
             y = VisibleHumidic_Like)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  theme(panel.grid = element_blank())

summary(lm(data = res_data, VisibleHumidic_Like ~ res))
```





### Mean Values
```{r}
Param_data_mean <- res_data_means %>% 
  relocate(spRichness, .after = res)

scaleParam <- as_tibble(scale(Param_data_mean[,6:29], scale = TRUE, center = TRUE)) %>% 
  bind_cols(res_data %>% select(res)) %>% 
  relocate(res, .before = spRichness)


Residuals_means_pval <- tibble(Parameter = as.character(),
                        pvalue = as.numeric(),
                        adj_r_squared = as.numeric(),
                        beta = as.numeric())

for(i in 2:ncol(scaleParam)){
  
  Parameter <- colnames(scaleParam)[i] # select dependent parameter
  
  mod <- lm(paste(Parameter, "~", "res"), data = scaleParam)
  pvalue <- summary(mod)[4]$coefficients[8]
  r_squared <- summary(mod)[8]$r.squared
  beta <- summary(mod)$coefficients[2]
  
  temp <- as_tibble(cbind(Parameter, pvalue, r_squared)) %>% 
    mutate(pvalue = as.numeric(pvalue),
           r_squared = as.numeric(r_squared),
           beta = as.numeric(beta))
  
  Residuals_means_pval <- Residuals_means_pval %>% 
    rbind(temp)
}


# view table of pvalues
Residuals_means_pval

# view bar graph of pvalues 
Residuals_means_pval %>% 
  arrange(beta) %>% 
  ggplot(aes(x = fct_reorder(Parameter, beta), y = beta)) +
  geom_col() +
  geom_hline(yintercept = 0.0) +
  theme_bw() +
  labs(y = "effect size: ~ Residuals from lm(Richness ~ Rugosity)",
       title = "Mean of biogeochemical values across Seasons") +
  theme(axis.text.x = element_text(angle = 90))
  
```

```{r}
ggplot(data = res_data_means,
       aes(x = del15N,
           y = res)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text_repel(aes(label = CowTagID)) +
  theme_bw() +
  theme(panel.grid = element_blank())

ggplot(data = res_data_means,
       aes(x = Ammonia_umolL,
           y = res)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text_repel(aes(label = CowTagID)) +
  theme_bw() +
  theme(panel.grid = element_blank())

ggplot(data = res_data_means,
       aes(x = VisibleHumidic_Like,
           y = res)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text_repel(aes(label = CowTagID)) +
  theme_bw() +
  theme(panel.grid = element_blank())
  
```




# nMDS


```{r}
# create levels for CowTagID order
CTlevels <- c('V1','V2','V3','V4','V5','V6','V7','V8','V9','V10','V11','V12','V14','V15','V16','V17','V18','V19','V20')
```

## nMDS Plot: Site characteristics (CV Values)

### Prep data

nMDS:
- create ordination output using bray curtis dissimilarity matrix
- remove non-numerical data
- distance = is the type of dissimilarity matrix
- k is number of dimensions (you can try different ones to look at different stress)
```{r}

richMDS <- res_data %>%
  select(CowTagID, # res, cannot use residuals because of negative values
         Salinity, Temperature, pH, 
         Phosphate_umolL, Silicate_umolL, NN_umolL, Ammonia_umolL,
         M_C, HIX, VisibleHumidic_Like, Tryptophan_Like, Tyrosine_Like
         #del15N, C_N, N_percent,  # remove because of mean values
         #meanRugosity # remove 
         ) # removed TA, substrate, distance to seep, and CT depth, and rugosity

richMDS$CowTagID <- factor(richMDS$CowTagID, levels = CTlevels)
richMDS <- richMDS %>% 
  arrange(CowTagID) %>% # arrange data in order of CowTagID for ordination plot later
  select(-CowTagID)

richOrd <- metaMDS(richMDS, k=2, distance='bray') 

# stress with k=2 dimensions. Is it < 0.3? 
richOrd$stress

# stress plot
# want to minimize scatter
stressplot(richOrd)
```

### Plot nMDS ordinations
```{r}

# get points for site characteristics
richRow <- rownames(richOrd$species) # get characteristic names

# with substrate
# richMDS1 <- c(richOrd$species[1:23]) # MDS1 for characteristics
# richMDS2 <- c(richOrd$species[24:46]) # MDS2 for characteristics

# without substrate
richMDS1 <- c(richOrd$species[,1]) # MDS1 for characteristics
richMDS2 <- c(richOrd$species[,2]) # MDS2 for characteristics

richData <- as_tibble(cbind(richRow, richMDS1, richMDS2)) %>%  # bind all cols into tibble
  mutate(richMDS1 = as.numeric(richMDS1), # as numeric 
         richMDS2 = as.numeric(richMDS2))
# get points for CowTagIDs
richRowb <- as.character(CTlevels) # assign CowTagID names
richMDS1b <- c(richOrd$points[,1]) # MDS1 for CowTagID
richMDS2b <- c(richOrd$points[,2]) # MDS2 for CowTagID
richDatab <- as_tibble(cbind(richRowb, richMDS1b, richMDS2b)) %>%  # bind all cols into tibble
  mutate(richMDS1b = as.numeric(richMDS1b), # as numeric 
         richMDS2b = as.numeric(richMDS2b)) %>% 
  rename(CowTagID = richRowb) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m))

# plotting
nMDSplot <- ggplot(data = richData, # sit characteristics
       aes(x = richMDS1,
           y = richMDS2)) +
  # geom_point(color = "black", 
  #            size = 3) +
  geom_point(data = richDatab, # CowTagIDs
             aes(x = richMDS1b,
                 y = richMDS2b,
                 fill = dist_to_seep_m),
             shape = 22,
             size = 4,
             color = "black") +
  geom_text(data = richData, # site characteristics
                  aes(x = richMDS1,
                      y = richMDS2,
                      label = richRow),
                  size = 3) +
  # geom_text_repel(data = richDatab, # CowTagIDs
  #                 aes(x = richMDS1b,
  #                     y = richMDS2b,
  #                     label = richRowb),
  #                 size = 3) +
  coord_cartesian(xlim = c(-0.12, 0.22), ylim = c(-0.18, 0.15)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "top") +
  labs(fill = "Distance from seep (m)")

nMDSplot
ggsave(here("Output","nMDS","nMDS_Varari_param_dist.png"), nMDSplot, width = 14, height = 10)
```


## nMDS Plot: Taxon groups (CV Values)

### Prep data
```{r}

speciesMDS <- comp %>%
  left_join(taxa) %>% 
  filter(Taxa != 'Hard Substrate', Taxa != 'Sand') %>% 
  select(CowTagID, Taxa, Taxon_Group, sp_percent) %>% 
  group_by(CowTagID, Taxon_Group) %>% 
  summarise(perc = sum(sp_percent)) %>% 
  pivot_wider(names_from = "Taxon_Group", values_from = "perc") %>% 
  mutate_at(vars('Anemone':'Cyanobacteria'), .funs = ~if_else(is.na(.)==TRUE, 0, .)) %>%  # put 0 %cover for all NA values
  ungroup()

speciesMDS$CowTagID <- factor(speciesMDS$CowTagID, levels = CTlevels)

speciesMDS <- speciesMDS %>% 
  arrange(CowTagID) %>% 
  select(-c(CowTagID))


speciesOrd <- metaMDS(speciesMDS, k=2, distance='bray') 

# stress with k=2 dimensions. Is it < 0.3? 
speciesOrd$stress

# stress plot
# want to minimize scatter
stressplot(speciesOrd)
```

### Plot nMDS ordinations
```{r}

# get points for species
speciesRow <- rownames(speciesOrd$species) # get characteristic names
speciesMDS1 <- c(speciesOrd$species[,1]) # MDS1 for characteristics
speciesMDS2 <- c(speciesOrd$species[,2]) # MDS2 for characteristics
speciesData <- as_tibble(cbind(speciesRow, speciesMDS1, speciesMDS2)) %>%  # bind all cols into tibble
  mutate(speciesMDS1 = as.numeric(speciesMDS1), # as numeric 
         speciesMDS2 = as.numeric(speciesMDS2)) %>% 
  rename(Taxon_Group = speciesRow) %>% 
  #mutate(Taxon_Group = if_else(Taxa == "Hard Substrate", "Abiotic", Taxon_Group)) %>% 
  select(speciesMDS1, speciesMDS2, Taxon_Group)
# get points for
speciesRowb <- as.character(CTlevels) # assign CowTagID
speciesMDS1b <- speciesOrd$points[,1] # MDS1 for CowTagID
speciesMDS2b <- speciesOrd$points[,2] # MDS2 for CowTagID
speciesDatab <- as_tibble(cbind(speciesRowb, speciesMDS1b, speciesMDS2b)) %>%  # bind all cols into tibble
  mutate(speciesMDS1b = as.numeric(speciesMDS1b), # as numeric
         speciesMDS2b = as.numeric(speciesMDS2b)) %>% 
  rename(CowTagID = speciesRowb) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))


```

### plot nMDS ~ distance  (CV Values)
```{r}

# plotting
nMDSplot_dist <- ggplot(data = speciesData, # site characteristics
       aes(x = speciesMDS1,
           y = speciesMDS2)) +
  #geom_point(size = 3) +
  geom_point(data = speciesDatab, # CowTagIDs
             aes(x = speciesMDS1b,
                 y = speciesMDS2b,
                 fill = dist_to_seep_m),
             shape = 22,
             size = 3) +
  geom_text(data = speciesData, # site characteristics
                  aes(x = speciesMDS1,
                      y = speciesMDS2,
                      label = speciesRow),
                  size = 3) +
  #coord_cartesian(xlim = c(-2, 2), ylim = c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "top") +
  labs(fill = "Distance from seep (m)")

nMDSplot_dist
ggsave(here("Output","nMDS","nMDS_Varari_taxa_dist.png"), nMDSplot_dist, width = 14, height = 10)
```



```{r}
speciesPermFull <- cbind(speciesMDS, CTlevels) %>%  # bind cowTagIDs
  rename(CowTagID = CTlevels) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel_dist<-adonis2(speciesPermFull[,-c(11:15)]~dist_to_seep_m, speciesPermFull, permutations = 999, 
                       method="bray") 
permanovamodel_dist


permanovamodel_am <- adonis2(speciesPermFull[,-c(11:15)]~Ammonia_umolL, speciesPermFull, permutations = 999, 
                       method="bray") 
permanovamodel_am


permanovamodel_VH <- adonis2(speciesPermFull[,-c(11:15)]~VisibleHumidic_Like, speciesPermFull, permutations = 999, 
                       method="bray")  
permanovamodel_VH


permanovamodel_15N <- adonis2(speciesPermFull[,-c(11:15)]~del15N, speciesPermFull, permutations = 999, 
                       method="bray") 
permanovamodel_15N

```




## nMDS Plot: Taxon groups (mean Values)

### plot nMDS ~ distance  (mean Values)
```{r}

speciesDatab_mean <- as_tibble(cbind(speciesRowb, speciesMDS1b, speciesMDS2b)) %>%  # bind all cols into tibble
  mutate(speciesMDS1b = as.numeric(speciesMDS1b), # as numeric
         speciesMDS2b = as.numeric(speciesMDS2b)) %>% 
  rename(CowTagID = speciesRowb) %>% 
  left_join(res_data_means %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))


# plotting
nMDSplot_dist_mean <- ggplot(data = speciesData, # site characteristics
       aes(x = speciesMDS1,
           y = speciesMDS2)) +
  #geom_point(size = 3) +
  geom_point(data = speciesDatab_mean, # CowTagIDs
             aes(x = speciesMDS1b,
                 y = speciesMDS2b,
                 fill = dist_to_seep_m),
             shape = 22,
             size = 3) +
  geom_text(data = speciesData, # site characteristics
                  aes(x = speciesMDS1,
                      y = speciesMDS2,
                      label = speciesRow),
                  size = 3) +
  #coord_cartesian(xlim = c(-2, 2), ylim = c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "top") +
  labs(fill = "Distance from seep (m)")

nMDSplot_dist_mean
ggsave(here("Output","nMDS","nMDS_Varari_taxa_dist_mean.png"), nMDSplot_dist_mean, width = 14, height = 10)
```


```{r}
speciesPermFull_mean <- cbind(speciesMDS, CTlevels) %>%  # bind cowTagIDs
  rename(CowTagID = CTlevels) %>% 
  left_join(res_data_means %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel_dist_mean<-adonis2(speciesPermFull_mean[,-c(11:15)]~dist_to_seep_m, speciesPermFull_mean, permutations = 999, 
                       method="bray") 
permanovamodel_dist_mean



permanovamodel_am_mean <- adonis2(speciesPermFull_mean[,-c(11:15)]~Ammonia_umolL, speciesPermFull_mean, permutations = 999, 
                       method="bray") 
permanovamodel_am_mean



permanovamodel_VH_mean <- adonis2(speciesPermFull_mean[,-c(11:15)]~VisibleHumidic_Like, speciesPermFull_mean, permutations = 999, 
                       method="bray")  
permanovamodel_VH_mean


permanovamodel_15N_mean <- adonis2(speciesPermFull_mean[,-c(11:15)]~del15N, speciesPermFull_mean, permutations = 999, 
                       method="bray") 
permanovamodel_15N_mean

```



# Various taxon groups nMDS

```{r}

# calcifiers 

calcifiersComp <- comp %>%
  left_join(taxa) %>% 
  filter(Taxa != 'Hard Substrate', Taxa != 'Sand') %>% 
  select(CowTagID, Taxa, Taxon_Group, sp_percent) %>% 
  mutate(Taxon_Group = if_else(Taxon_Group == "Coral", "Calcifiers",
                       if_else(Taxon_Group == "CCA", "Calcifiers", Taxon_Group)))


speciesMDS_calc <- calcifiersComp %>% 
  group_by(CowTagID, Taxon_Group) %>% 
  summarise(perc = sum(sp_percent)) %>% 
  pivot_wider(names_from = "Taxon_Group", values_from = "perc") %>% 
  mutate_at(vars('Anemone':'Cyanobacteria'), .funs = ~if_else(is.na(.)==TRUE, 0, .)) %>%  # put 0 %cover for all NA values
  ungroup()

speciesMDS_calc$CowTagID <- factor(speciesMDS_calc$CowTagID, levels = CTlevels)

speciesMDS_calc <- speciesMDS_calc %>% 
  arrange(CowTagID) %>% 
  select(-c(CowTagID))

speciesOrd_calc <- metaMDS(speciesMDS_calc, k=2, distance='bray') 

speciesOrd_calc$stress

stressplot(speciesOrd_calc)


## macroalgae and calcifiers

calcMacroComp <- comp %>%
  left_join(taxa) %>% 
  filter(Taxa != 'Hard Substrate', Taxa != 'Sand') %>% 
  select(CowTagID, Taxa, Taxon_Group, sp_percent) %>% 
  mutate(Taxon_Group = if_else(Taxon_Group == "Coral", "Calcifiers",
                       if_else(Taxon_Group == "CCA", "Calcifiers",
                       if_else(Taxon_Group == "Rhodophyta", "Macroalgae",
                       if_else(Taxon_Group == "Phaeophyta", "Macroalgae",
                       if_else(Taxon_Group == "Chlorophyta", "Macroalgae", Taxon_Group))))))

speciesMDS_calcMacro <- calcMacroComp %>% 
  group_by(CowTagID, Taxon_Group) %>% 
  summarise(perc = sum(sp_percent)) %>% 
  pivot_wider(names_from = "Taxon_Group", values_from = "perc") %>% 
  mutate_at(vars('Anemone':'Cyanobacteria'), .funs = ~if_else(is.na(.)==TRUE, 0, .)) %>%  # put 0 %cover for all NA values
  ungroup()

speciesMDS_calcMacro$CowTagID <- factor(speciesMDS_calcMacro$CowTagID, levels = CTlevels)

speciesMDS_calcMacro <- speciesMDS_calcMacro %>% 
  arrange(CowTagID) %>% 
  select(-c(CowTagID))

speciesOrd_calcMacro <- metaMDS(speciesMDS_calcMacro, k=2, distance='bray') 

# stress with k=2 dimensions. Is it < 0.3? 
speciesOrd_calcMacro$stress

# stress plot
# want to minimize scatter
stressplot(speciesOrd_calcMacro)
```
## calcifiers
### Plot nMDS ordinations
```{r}

# get points for species
speciesRow <- rownames(speciesOrd_calc$species) # get characteristic names
speciesMDS1 <- c(speciesOrd_calc$species[,1]) # MDS1 for characteristics
speciesMDS2 <- c(speciesOrd_calc$species[,2]) # MDS2 for characteristics
speciesData <- as_tibble(cbind(speciesRow, speciesMDS1, speciesMDS2)) %>%  # bind all cols into tibble
  mutate(speciesMDS1 = as.numeric(speciesMDS1), # as numeric 
         speciesMDS2 = as.numeric(speciesMDS2)) %>% 
  rename(Taxon_Group = speciesRow) %>% 
  #mutate(Taxon_Group = if_else(Taxa == "Hard Substrate", "Abiotic", Taxon_Group)) %>% 
  select(speciesMDS1, speciesMDS2, Taxon_Group)
# get points for
speciesRowb <- as.character(CTlevels) # assign CowTagID
speciesMDS1b <- speciesOrd_calc$points[,1] # MDS1 for CowTagID
speciesMDS2b <- speciesOrd_calc$points[,2] # MDS2 for CowTagID
speciesDatab <- as_tibble(cbind(speciesRowb, speciesMDS1b, speciesMDS2b)) %>%  # bind all cols into tibble
  mutate(speciesMDS1b = as.numeric(speciesMDS1b), # as numeric
         speciesMDS2b = as.numeric(speciesMDS2b)) %>% 
  rename(CowTagID = speciesRowb) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))


# plotting
nMDSplot_dist_calc <- ggplot(data = speciesData, # site characteristics
       aes(x = speciesMDS1,
           y = speciesMDS2)) +
  #geom_point(size = 3) +
  geom_point(data = speciesDatab, # CowTagIDs
             aes(x = speciesMDS1b,
                 y = speciesMDS2b,
                 fill = dist_to_seep_m),
             shape = 22,
             size = 3) +
  geom_text(data = speciesData, # site characteristics
                  aes(x = speciesMDS1,
                      y = speciesMDS2,
                      label = speciesRow),
                  size = 3) +
  #coord_cartesian(xlim = c(-2, 2), ylim = c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "top") +
  labs(fill = "Distance from seep (m)")

nMDSplot_dist_calc
#ggsave(here("Output","nMDS","nMDS_Varari_taxa_dist.png"), nMDSplot_dist, width = 14, height = 10)
```

```{r}
speciesPermFull_calc <- cbind(speciesMDS_calc, CTlevels) %>%  # bind cowTagIDs
  rename(CowTagID = CTlevels) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel_dist<-adonis2(speciesPermFull_calc[,-c(10:14)]~dist_to_seep_m, speciesPermFull_calc, permutations = 999, 
                       method="bray") 
permanovamodel_dist


permanovamodel_am <- adonis2(speciesPermFull_calc[,-c(10:14)]~Ammonia_umolL, speciesPermFull_calc, permutations = 999, 
                       method="bray") 
permanovamodel_am


permanovamodel_VH <- adonis2(speciesPermFull_calc[,-c(10:14)]~VisibleHumidic_Like, speciesPermFull_calc, permutations = 999, 
                       method="bray")  
permanovamodel_VH


permanovamodel_15N <- adonis2(speciesPermFull_calc[,-c(10:14)]~del15N, speciesPermFull_calc, permutations = 999, 
                       method="bray") 
permanovamodel_15N

```


## calcifiers and macroalgae
### Plot nMDS ordinations
```{r}

# get points for species
speciesRow <- rownames(speciesOrd_calcMacro$species) # get characteristic names
speciesMDS1 <- c(speciesOrd_calcMacro$species[,1]) # MDS1 for characteristics
speciesMDS2 <- c(speciesOrd_calcMacro$species[,2]) # MDS2 for characteristics
speciesData <- as_tibble(cbind(speciesRow, speciesMDS1, speciesMDS2)) %>%  # bind all cols into tibble
  mutate(speciesMDS1 = as.numeric(speciesMDS1), # as numeric 
         speciesMDS2 = as.numeric(speciesMDS2)) %>% 
  rename(Taxon_Group = speciesRow) %>% 
  #mutate(Taxon_Group = if_else(Taxa == "Hard Substrate", "Abiotic", Taxon_Group)) %>% 
  select(speciesMDS1, speciesMDS2, Taxon_Group)
# get points for
speciesRowb <- as.character(CTlevels) # assign CowTagID
speciesMDS1b <- speciesOrd_calcMacro$points[,1] # MDS1 for CowTagID
speciesMDS2b <- speciesOrd_calcMacro$points[,2] # MDS2 for CowTagID
speciesDatab <- as_tibble(cbind(speciesRowb, speciesMDS1b, speciesMDS2b)) %>%  # bind all cols into tibble
  mutate(speciesMDS1b = as.numeric(speciesMDS1b), # as numeric
         speciesMDS2b = as.numeric(speciesMDS2b)) %>% 
  rename(CowTagID = speciesRowb) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))


# plotting
nMDSplot_dist_calcMacro <- ggplot(data = speciesData, # site characteristics
       aes(x = speciesMDS1,
           y = speciesMDS2)) +
  #geom_point(size = 3) +
  geom_point(data = speciesDatab, # CowTagIDs
             aes(x = speciesMDS1b,
                 y = speciesMDS2b,
                 fill = dist_to_seep_m),
             shape = 22,
             size = 3) +
  geom_text(data = speciesData, # site characteristics
                  aes(x = speciesMDS1,
                      y = speciesMDS2,
                      label = speciesRow),
                  size = 3) +
  #coord_cartesian(xlim = c(-2, 2), ylim = c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "top") +
  labs(fill = "Distance from seep (m)")

nMDSplot_dist_calcMacro
#ggsave(here("Output","nMDS","nMDS_Varari_taxa_dist.png"), nMDSplot_dist, width = 14, height = 10)
```

```{r}
speciesPermFull_calcMacro <- cbind(speciesMDS_calcMacro, CTlevels) %>%  # bind cowTagIDs
  rename(CowTagID = CTlevels) %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel_dist<-adonis2(speciesPermFull_calcMacro[,-c(8:12)]~dist_to_seep_m, speciesPermFull_calcMacro, permutations = 999, 
                       method="bray") 
permanovamodel_dist # p = 0.086


permanovamodel_am <- adonis2(speciesPermFull_calcMacro[,-c(8:12)]~Ammonia_umolL, speciesPermFull_calcMacro, permutations = 999, 
                       method="bray") 
permanovamodel_am


permanovamodel_VH <- adonis2(speciesPermFull_calcMacro[,-c(8:12)]~VisibleHumidic_Like, speciesPermFull_calcMacro, permutations = 999, 
                       method="bray")  
permanovamodel_VH


permanovamodel_15N <- adonis2(speciesPermFull_calcMacro[,-c(8:12)]~del15N, speciesPermFull_calcMacro, permutations = 999, 
                       method="bray") 
permanovamodel_15N

```









# All species nMDS

## ~ distance
```{r}
speciesMDS2 <- comp %>%
  left_join(taxa) %>% 
  filter(Taxa != 'Hard Substrate', Taxa != 'Sand') %>% 
  select(CowTagID, Taxa,sp_percent) %>% 
  group_by(CowTagID, Taxa) %>% 
  summarise(perc = sum(sp_percent)) %>% 
  pivot_wider(names_from = "Taxa", values_from = "perc") %>% 
  mutate_at(vars('Crustose Corallines':'Valonia macrophysa'), .funs = ~if_else(is.na(.)==TRUE, 0, .)) %>%  # put 0 %cover for all NA values
  ungroup()

speciesMDS2$CowTagID <- factor(speciesMDS2$CowTagID, levels = CTlevels)

speciesPermFull2 <- speciesMDS2 %>% 
  left_join(res_data %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel2<-adonis2(speciesPermFull2[,-c(1,60:63)]~VisibleHumidic_Like, speciesPermFull2, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2 


permanovamodel2<-adonis2(speciesPermFull2[,-c(1,60:63)]~dist_to_seep_m, speciesPermFull2, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2 # *

permanovamodel2<-adonis2(speciesPermFull2[,-c(1,60:63)]~Ammonia_umolL, speciesPermFull2, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2 

permanovamodel2<-adonis2(speciesPermFull2[,-c(1,60:63)]~del15N, speciesPermFull2, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2 
```


## Mean values
```{r}
speciesPermFull2_means <- speciesMDS2 %>% 
  left_join(res_data_means %>% select(CowTagID, dist_to_seep_m, VisibleHumidic_Like, Ammonia_umolL, del15N))

permanovamodel2_means<-adonis2(speciesPermFull2_means[,-c(1,60:63)]~VisibleHumidic_Like, speciesPermFull2_means, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2_means 


permanovamodel2_means<-adonis2(speciesPermFull2_means[,-c(1,60:63)]~dist_to_seep_m, speciesPermFull2_means, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2_means # *

permanovamodel2_means<-adonis2(speciesPermFull2_means[,-c(1,60:63)]~Ammonia_umolL, speciesPermFull2_means, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2_means 

permanovamodel2_means<-adonis2(speciesPermFull2_means[,-c(1,60:63)]~del15N, speciesPermFull2_means, permutations = 999, 
                       method="bray") # should change out cowtagid with some grouping name
permanovamodel2_means 
```



### Plot with more broad taxon groups

```{r, echo = F, eval = F}
# create my own palette
taxonPalette <- tibble('Taxon_Group' = as.character(c("Abiotic", "Turf", "Macroalgae", "Calcifier", "Sponge", "Corallimorph", "Anemone", "Cyanobacteria")),
                       'PalColor' = as.character(c("black", "#00496F", "#086B8B", "#2E9093", "#EDC132", "#ED950A", "#DD4124", "#b99095")))

# broad taxonomic groups
taxonData <- speciesData %>% 
  mutate(Taxon_Group = if_else(Taxon_Group == "Chlorophyta", "Macroalgae", 
                       if_else(Taxon_Group == "Rhodophyta", "Macroalgae",
                       if_else(Taxon_Group == "Phaeophyta", "Macroalgae",
                       if_else(Taxon_Group == "Coral", "Calcifier", 
                       if_else(Taxon_Group == "CCA", "Calcifier",
                       if_else(Taxon_Group == "Sand", "Abiotic", Taxon_Group))))))) %>% 
  left_join(taxonPalette)

Pal <- as.character(taxonData$PalColor)
names(Pal) <- as.character(taxonData$Taxon_Group)


# plotting
nMDSplot2b <- ggplot(data = taxonData, # site characteristics
       aes(x = speciesMDS1,
           y = speciesMDS2)) +
  geom_point(aes(color = Taxon_Group),
             size = 3) +
  geom_point(data = speciesDatab, # CowTagIDs
             aes(x = speciesMDS1b,
                 y = speciesMDS2b),
             shape = 22,
             size = 3) +
  geom_text_repel(data = taxonData, # site characteristics
                  aes(x = speciesMDS1,
                      y = speciesMDS2,
                      label = speciesRow),
                  size = 2,
                  max.overlaps = 15) +
  #coord_cartesian(xlim = c(-2, 2), ylim = c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  #scale_color_manual(values = Pal) +
  labs(x = "MDS1", y = "MDS2", color = "Taxonomic Group")

nMDSplot2b
#ggsave(here("Output","nMDS_Varari.png"), nMDSplot, width = 14, height = 9)
```



### Parameter regression pvalues without seep and V13
```{r, eval = F}
red_Full <- res_data %>% 
  select(-c(Location, lat, lon))

Param_pval <- tibble(Parameter_Ind = as.character(),
                         Parameter_Dep = as.character(),
                        pvalue = as.numeric(),
                        adj_r_squared = as.numeric())


for(i in 2:ncol(red_Full)) {
  
  Parameter_Ind <- colnames(red_Full)[i] # select ith parameter as x axis variable 
  
  temp1 <- tibble(Parameter_Ind = as.character(),
                         Parameter_Dep = as.character(),
                        pvalue = as.numeric(),
                        adj_r_squared = as.numeric()) 
  
  for(j in 3:ncol(red_Full)) {
  
  if(i != j){
    Parameter_Dep <- colnames(red_Full)[j] # select ith parameter as x axis variable
  } else if(j < 25) {
    Parameter_Dep <- colnames(red_Full)[(j+1)] # select ith parameter as x axis variable
  } else if(j == 25) {
    Parameter_Dep <- colnames(red_Full)[2] # select ith parameter as x axis variable
  }
    
  mod <- lm(paste(Parameter_Ind, "~", Parameter_Dep), data = red_Full)
  pvalue <- summary(mod)[4]$coefficients[8]
  adj_r_squared <- summary(mod)[9]$adj.r.squared
  
  temp <- as_tibble(cbind(Parameter_Ind, Parameter_Dep, pvalue, adj_r_squared)) %>% 
    mutate(pvalue = as.numeric(pvalue),
           adj_r_squared = as.numeric(adj_r_squared))
  
  temp1 <- temp1 %>% 
    rbind(temp)
  
  }
  
    Param_pval <- Param_pval %>% 
    rbind(temp1) %>% 
      distinct()
}

# view table of pvalues
Param_pval
write_csv(Param_pval, here("Data", "biogeochem", "biogeochem_pvalus_res_noSeepV13.csv"))

```










